### 2021-09-13 plos_latex_template.tex
finished 1st round of comments - sorry for the delay - Batool

And most importantly, take care of your team and their well-being! Check on them regularly and make sure everyone is comfortable. This represents a great amount of work and you'll need to support each other in the long run.

@@

Make sure that every selection committee --the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees --are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has an implicit biases, to recognize them, and try to counteract them. An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles. The regional and local communities in your field are also good sources to tap into. 
For example, for useR! 2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask their needs and barriers, and explicitly invite them to contribute to the conference.

@@

In international conferences, English is often the official language. Submissions, presentations, tutorials, and workshops are in English. The platforms, the webpage, and official communications are also in English. While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, opening up the conference to other languages could make it less intimidating to people who are not fluent in English \cite{ninerBetterWhomLeveling2021}. Excluding them may potentially lead to missing innovative contributions due to a language barrier. Advertising the conference in several languages and considering having non-English workshops and presentations (with or without captions in English) could help overcome this barrier. For instance, hosting one international day/session per conference might be a great place to start!

@@

The ten rules stated here can be adapted depending on the conference format and settings.
When organizing useR!2021, we engaged in most of the practices mentioned here, and learned others along the way, so we share them here as part of our learning process. 
We organized useR! during a global pandemic, and as a team, this was a challenging journey. 
Very early on, we tried to recognize that our resources were limited (both in time of the organizing team and financially). 
We had to let go of certain aspects that were important to many of us. 
From that experience, this is our last message: your conference may not become perfectly inclusive and accessible, but the changes you make will make a difference.
Also, keep working to improve inclusion in the next edition of the conference! to document and balance your work, design surveys, ask for feedback from attendees and organizers. Transfer your constructed knowledge to the next organizing team, and adapt year-to-year to include more dimensions of diversity, to address technical and social changes, and to include insights from the new organizing team.
If more conferences and domains apply these rules, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
And you will make a change towards healthier, stronger, and more inclusive communities.

>>>>

And most importantly, take care of your team and prioritize their well-being! Check on them regularly and make sure everyone is comfortable. This represents a great amount of work and you'll need to support each other in the long run.

@@

Make sure that every selection committee --the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees --are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has an implicit biases, to recognize them, and try to counteract them.  An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles. The regional and local communities in your field are also good sources to tap into. 
For example, for useR! 2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask their needs and barriers, and explicitly invite them to contribute to the conference.Since many conferences are held with the primary goal of displaying cutting-edge work led by senior scientists and academics, early career researchers are often disadvantaged. Diversity should spread among all speaking roles, and the conference should also help raise the profiles of early career researchers, fostering collaboration and building their skills.

@@

In international conferences, English is often the official language. Submissions, presentations, tutorials, and workshops are in English. The platforms, the webpage, and official communications are also in English. While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, opening up the conference to other languages could make it less intimidating to people who are not fluent in English \cite{ninerBetterWhomLeveling2021}. Excluding them may potentially lead to missing innovative contributions due to a language barrier.

Meeting the need of linguistically diverse audience is an important aspect in inclusion that is often overlooked. Non-native English speakers may feel intimidated to ask questions or raise their views. Providing a welcoming and diverse environment by encouraging the full participation of those individuals is critical (Rule 2). If the panellist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language. Consider letting speakers know in advance that they can ask for a question to be translated to them in their native language and that they can respond in their native language.

Advertising the conference in several languages and considering having non-English workshops and presentations (with or without captions in English) could help overcome this barrier. For instance, hosting one international day/session per conference might be a great place to start!

@@

The ten rules stated here can be adapted depending on the conference format and settings.
When organizing useR!2021, we engaged in most of the practices mentioned here, and learned others along the way, so we share them here as part of our learning process. 
We organized useR! during a global pandemic, and as a team, this was a challenging journey. 
Very early on, we tried to recognize that our resources were limited (both in time of the organizing team and financially). 
We had to let go of certain aspects that were important to many of us. 
From that experience, this is our last message: your conference may not become perfectly inclusive and accessible, but the changes you make will make a difference.
Also, keep working to improve inclusion in the next edition of the conference! to document and balance your work, design surveys, ask for feedback from attendees and organizers. Transfer your constructed knowledge to the next organizing team, and adapt year-to-year to include more dimensions of diversity, to address technical and social changes, and to include insights from the new organizing team.  Always remember inclusion is an ambition that organisers can and should continually pursue by allowing room for learning and improvement.
If more conferences and domains apply these rules, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead.
And you will make a change towards healthier, stronger, and more inclusive communities.
